---
title: "FML Final Project"
author: "Nikeeta Rani"
date: "2024-04-16"
output:
  html_document: default
  pdf_document: default
---

# Loading Necessary Libraries 
```{r}
library(readxl)
library(dplyr)
library(ISLR)
library(caret)
library(corrplot)
library(GGally)
library(class)
library(gmodels)
library(e1071)
library(pROC)
library(psych)
library(FactoMineR)
library(missMDA)
library(devtools)
library(ggbiplot)
library(factoextra)
library(ggplot2)
library(tidyr)
```

# Loading the Data 
```{r}
Credit_Raw<- read_excel("C:/Users/hp/Desktop/KSU/FML/Final Project/CreditScoreData.xlsx")
```

# Removing one level of Response i.e "Standard"
```{r}
Credit<-Credit_Raw[Credit_Raw$Credit_Score != "Standard", ]
```

```{r}
head(Credit)
```

# Descriptive Statistics
```{r}
summary(Credit)
```

```{r}
sapply(Credit, class)
```

```{r}
colnames(Credit)
```
# Identifying Missing Values 
```{r}
Missing_Data<-colMeans(is.na(Credit))
Missing_Data
```
### Analysis: Type of Loan column has 9% of the missing data 

# Removing Missing Values 
```{r}
Credit<-na.omit(Credit)
Missing_Data<-colMeans(is.na(Credit))
Missing_Data
```

# Removing Unnecassary Variables (ID, Customer ID, Month, Name and SSN) and Few Categorical Variables 
```{r}
Credit_Score_Filtered <- Credit[, -c(1:4, 6,7,14,26)]
colnames(Credit_Score_Filtered)
```

```{r}
summary(Credit_Score_Filtered)
```

# BoxPlot 

### Identifying Numeric Variables
```{r}
numeric_df <- select_if(Credit_Score_Filtered, is.numeric)
```

### Box Plot1 
```{r}
Plot_Box1 <- data.frame(
  Num_Bank_Accounts = numeric_df$Num_Bank_Accounts, 
  Num_Credit_Card = numeric_df$Num_Credit_Card, 
  Num_of_Loan = numeric_df$Num_of_Loan
)
par(mar = c(7, 5, 4, 2) + 0.1) 
boxplot(Plot_Box1, las = 2, cex.axis = 0.7)
```
### Boxplot 2
```{r}
Plot_Box2 <- data.frame(
  Interest_Rate = numeric_df$Interest_Rate, 
  Delay_From_Due_Date = numeric_df$Delay_from_due_date,
  Changed_Credit_Limit= numeric_df$Changed_Credit_Limit,
  Credit_Utilization_Ratio = numeric_df$Credit_Utilization_Ratio,
  Age = numeric_df$Age
)
par(mar = c(7, 5, 4, 2) + 0.1) 
boxplot(Plot_Box2, las = 2, cex.axis = 0.7)
```
### Boxplot 3
```{r}
Plot_Box3 <- data.frame(
  Num_Delayed_Payment=  numeric_df$Num_of_Delayed_Payment, 
  Number_Credit_Inquiries = numeric_df$Num_Credit_Inquiries,
  Outstanding_Debt = numeric_df$Outstanding_Debt
)
par(mar = c(7, 5, 4, 2) + 0.1) 
boxplot(Plot_Box3, las = 2, cex.axis = 0.7)
```

###  Boxplot 4 
```{r}
Plot_Box4 <- data.frame(
  Amount_Invested_Monthly=  numeric_df$Amount_invested_monthly, 
  Monthly_Balance = numeric_df$Monthly_Balance,
  Credit_History_Age_Months = numeric_df$Credit_History_Age_Months
)
par(mar = c(7, 5, 4, 2) + 0.1) 
boxplot(Plot_Box4, las = 2, cex.axis = 0.7)
```

### Box Plot 5,6 and 7 
```{r}
Boxplot_5 <- boxplot(numeric_df$Amount_invested_monthly)
Boxplot_6 <- boxplot(numeric_df$Monthly_Balance)
Boxplot_7 <- boxplot(numeric_df$Annual_Income)
```
```{r}
numeric_df <- select_if(Credit_Score_Filtered, is.numeric)
```


# Plotting Histrogram
```{r}
Hist_data <- numeric_df %>%
  gather(key = "Variable", value = "Value")

ggplot(Hist_data, aes(x = Value)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 30) +  # Set fill and outline color
  facet_wrap(~ Variable, scales = 'free') +
  theme_classic() +
  theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 
```

```{r}
ncol(Credit_Score_Filtered)
```

```{r}
summary(Credit_Score_Filtered)
```

# Transforming Categorical into Dummy Variables
```{r}
Credit_Score_Filtered$Credit_Score <- as.factor(Credit_Score_Filtered$Credit_Score)
levels(Credit_Score_Filtered$Credit_Score) <- c(Poor = "0", Good = "1")
unique(Credit_Score_Filtered$Credit_Score)
```
```{r}
Credit_Score_Filtered$Credit_Mix <- as.factor(Credit_Score_Filtered$Credit_Mix)
levels(Credit_Score_Filtered$Credit_Mix) <- c(Poor = "3", Good = "1", standard="2")
unique(Credit_Score_Filtered$Credit_Mix)
```
```{r}
Credit_Score_Filtered$Payment_of_Min_Amount <- as.factor(Credit_Score_Filtered$Payment_of_Min_Amount)
levels(Credit_Score_Filtered$Payment_of_Min_Amount) <- c(Yes = "1", No = "0")
unique(Credit_Score_Filtered$Payment_of_Min_Amount)
```

```{r}
Credit_Score_Filtered$Credit_Score <- as.numeric(as.character(Credit_Score_Filtered$Credit_Score))
Credit_Score_Filtered$Payment_of_Min_Amount <- as.numeric(as.character(Credit_Score_Filtered$Payment_of_Min_Amount))
Credit_Score_Filtered$Credit_Mix <- as.numeric(as.character(Credit_Score_Filtered$Credit_Mix))
```

```{r}
summary(Credit_Score_Filtered)
```

```{r}
ncol(Credit_Score_Filtered)
```

# Normalizing Data Set Except for out response i.e Credit Score
```{r}
Credit_Score_work <- Credit_Score_Filtered[, -20]
Normalized_Model <- preProcess(Credit_Score_work, method = c('center'))
Credit_Normalized_Excluded <- predict(Normalized_Model, Credit_Score_work)
Credit_Normalized <- cbind(Credit_Normalized_Excluded, Credit_Score = Credit_Score_Filtered[, 20])
summary(Credit_Normalized)
```

# Correlation Matrix
```{r}
Credit_Cor <- Credit_Normalized
correlation_matrix <- cor(Credit_Cor)
corrplot(correlation_matrix, method="circle")
```


# Selecting Features using PCA 

## PCA 1
```{r}
pc <- prcomp(Credit_Normalized,
             center = TRUE,
            scale. = TRUE)
```

`
# Pait Matrix -2 - With PCA 
```{r}
 pairs.panels(pc$x,
             gap=0,
             bg = c("red", "yellow", "blue")[Credit_Normalized$Credit_Score],
             pch=21)
```


```{r}
fviz_pca_var(pc, col.var = "cos2", 
             gradient.cols = c ("blue","orange", "green"),
             repel = TRUE)
             
```

```{r}
colnames(Credit_Normalized)
```
```{r}
Credit_Noramlized_After_PCA <- Credit_Normalized[, -c(1,2,4,5,6,7,9,10,11,12,14,17,18,19)]
colnames(Credit_Noramlized_After_PCA)
```
```{r}
  pc1 <- prcomp(Credit_Noramlized_After_PCA,
               center = TRUE,
              scale. = TRUE)
```

# Pair Matrix - 3 - With PCA
```{r}
pairs.panels(pc1$x,
             gap=0,
             bg = c("red", "yellow", "blue")[Credit_Noramlized_After_PCA$Credit_Score],
             pch=21)

```


## PCA 2 
```{r}
fviz_pca_var(pc1, col.var = "cos2", 
             gradient.cols = c ("blue","orange", "green"),
             repel = TRUE)
```

# Correlaton with top Fetaures
```{r}
Credit_Cor <- Credit_Noramlized_After_PCA
correlation_matrix <- cor(Credit_Cor)
corrplot(correlation_matrix, method="circle")
```


```{r}
ncol(Credit_Noramlized_After_PCA)
head(Credit_Noramlized_After_PCA)
```

# Partitioning the data in Testing(70%) and Training (30%) 
```{r}
set.seed(123)
Index_train <- createDataPartition(Credit_Noramlized_After_PCA$Credit_Score, p = 0.7, list = FALSE)
Train <- Credit_Noramlized_After_PCA[Index_train, ]
Test <- Credit_Noramlized_After_PCA[-Index_train, ]
```


```{r}
Train_Predictors <- Train[, 1:5]
Test_Predictors <- Test[, 1:5]
Train_labels <- Train[, 6]
Test_labels <- Test[, 6 ]
```

# Using Knn Method 

```{r}
Search_Grid <- expand.grid(k = 1:10) 
Train_labels <- as.factor(Train_labels)
model <- train(x = Train_Predictors, y = Train_labels, method = "knn", tuneGrid = Search_Grid) 
model 
```

```{r}
Predicted_Model <- knn(Train_Predictors, Test_Predictors, cl = Train_labels, k = 1)
head(Predicted_Model)
```

# Cross Table
```{r}
CrossTable(x=Test_labels,y=Predicted_Model,prop.chisq=FALSE )
```


# Confusion Matrix 
```{r}
Confusion_Matrix <- as.factor(Predicted_Model)
Test_labels <- as.factor(Test_labels)
Confusion_Matrix1 <- confusionMatrix(data = Predicted_Model, reference = Test_labels)
Confusion_Matrix1
```
# Confusion Matrix Using Knn - Recall, Precision and Accuracy 
```{r}
TN <- 784
FP <- 135
FN <- 159
TP <- 1909

Recall <- TP / (TP + FN)
Precision <- TP / (TP + FP)
Accuracy <- (TP + TN) / (TP + TN + FP + FN)

Accuracy
Recall
Precision
```


# Using Naive Bayes Model
```{r}
nb_Model <- naiveBayes(Credit_Score~., data = Train)
head(nb_Model)
```

```{r}
Predicted_Model_2<-predict(nb_Model, Test, type = "raw")
head(Predicted_Model_2, n=20)
```
# ROC curve
```{r}
roc_curve <- roc(Test$Credit_Score, Predicted_Model_2[, "1"])
plot(roc_curve, main = "ROC Curve", col = "blue")
auc_value <- auc(roc_curve)
auc_value
```

```{r}
Predicted_Model_2<-predict(nb_Model, Test)
head(Predicted_Model_2, n=6)
```

```{r}
levels(Predicted_Model)
levels(Predicted_Model_2)
```


```{r}
CrossTable(x=Test$Credit_Score,y=Predicted_Model_2,prop.chisq=FALSE )
```



# Confusion Matrix Using Naive Bayes
```{r}
Confusion_Matrix_nb <- as.factor(Predicted_Model_2)
Confusion_Matrix2 <- confusionMatrix(data = Confusion_Matrix_nb, reference = Test_labels)
Confusion_Matrix2
```
# Confusion Matrix Using Naive Bayes - Recall, Precision and Accuracy 
```{r}
TN <- 807
FP <- 112
FN <- 341
TP <- 1727

Recall <- TP / (TP + FN)
Precision <- TP / (TP + FP)
Accuracy <- (TP + TN) / (TP + TN + FP + FN)

Accuracy
Recall
Precision
```
```

  



















